{Investigating the Impact of Macroeconomic Indicators on the FTSE 250 Index Using Statistical and Machine Learning Methods}

### 2024-08-26 tex/paper.tex
revisions

\title{{Investigating the Impact of Macroeconomic Indicators
on the FTSE 250 Index Using Statistical and Machine Learning Methods}}
\author{
  \textbf{Mateen Khan} \\
  Mathematics Institute, University of Warwick  \\ 
  \\
  \textbf{Supervisor: Dr. Randa Herzallah} \\
  Mathematics Institute, University of Warwick
}
\begin{document}

@@

\begin{abstract}
    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate, and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
    In addition, it assesses the effectiveness of radial basis function neural networks (RBFNN) as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables. This study finds quantitative estimates for the short-term and long-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index. It also demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index.
    \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, radial basis function neural network, time-series forecasting}
\end{abstract}

@@

The stock market is a significant component of the financial system and 
plays a crucial role in the broader economy. Consequently, economic policy 
makers take the stock market into account when making fiscal policy 
decisions using a combination of quantitative and qualitative methods. 
These methods such as econometric modelling and forecasting, and empirical 
analysis allow them to understand the short-term and long-term effects of 
macroeconomic indicators
on the wider economy through the stock market (\citeboth{DemirgucKunt1996}).

@@

The significance of this study is to quantify the effects of macroeconomic indicators on the stock market as well as assessing the use of these indicators to forecast stock market prices over the short-term and long-term. In addition, the study seeks to understand which macroeconomic indicators are most relevant in capturing the dynamics of a stock market to aid forecasting.

@@

In order to analyse the dynamic relationship between the variables in the time series data, this study
will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error
Correction Model (ECM). This research also employs a radial basis function neural network to predict the price movements of the FTSE250.

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space and explains the rationale behind the selection of macroeconomic variables.
Section \ref{sec:meth} presents the methodology of this study. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
the study.

@@

Radial basis function neural networks are a type of artificial 
neural network that use radial basis functions at each node in the 
activation layer of the network.
The advantage of radial basis function neural networks lies in their ability to 
operate effectively in volatile financial time-series 
environments (\citeboth{cafferata2019}), their strong ability to 
generalise on such data (\citeboth{sharkawy2020}) as well as their efficacy on small datasets (\citeboth{kosarac2022}). 
\citeboth{esfandyari2016}
supports this, using a sample size of 70 to forecast daily NASDAQ 
Index values using an Artificial Neural Network; they achieved a minimum
test $R^2$ of 83$\%$.

Radial basis function neural networks have been used considerably for stock market forecasting. \citeboth{cao2004} 
used a radial basis function network based on the Optimal Partition Algorithm (OPA) to 
forecast S$\&$P 500 prices. \citeboth{dass2019} use a radial basis function neural network, utilising the Back-propagation 
Algorithm (BPA) instead to tune parameters; their results demonstrate 
that the radial basis function network outperformed a standard feed-forward neural network. 
\citeboth{ji2014} also found that the radial basis function network was able to predict the value 
of the Shanghai Composite Index on a 45 day time-horizon using the closing price, with a mean prediction error of 
1.4387$\%$. Recently, \citeboth{abotaleb2024}
used quarterly data from 2001 to 2023 to model the influence of the economic 
uncertainty index on the quarterly prices of various stock markets including the 
FTSE 100 Index using a radial basis function. They found that the model 
achieved a relative error of 27.2$\%$ in the case of the FTSE 100 Index.

Whilst \citeboth{abotaleb2024}, used quarterly data and a relatively minor economic indicator, the effectiveness of radial basis function neural networks in stock market forecasting using 
major macroeconomic indicators and coarse-grain time-series has not yet been investigated and presents an area for further 
research.

@@

Over the period covered by the dataset, the FTSE250 has 
shown substantial growth, reflecting the structural transition of the UK economy.
It has also experienced significant volatility, highlighting significant market shifts driven by 
various crises such as the Dotcom bubble burst, and geopolitical shocks.

@@

The USD/GBP exchange rate has exhibited moderate variability, exhibiting the stable relations between 
the two countries. Nonetheless, geopolitical shocks such as the 2008 financial crisis and Brexit both greatly affected the value of the Pound, 
which accounts for the large range.

The CPI, with an average of 88.36, has 
experienced significant fluctuations, highlighting periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
inflation such as the period between the early to late 2010s.

Interest rates have ranged from $0.1\%$ to $8.4\%$ with a standard deviation of $2.52\%$, reflecting the 
increasing role played by monetary policy in the UK since May 1997. This was when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}). The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
post-COVID inflationary pressures.

The M3 money supply has shown substantial variability, ranging from 494,181 to 3,704,873, 
in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
in response to the recessions caused by the 2008 financial crisis and the COVID-19 pandemic. These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.

@@

The study used an autoregressive distributed lag (ARDL) model in conjunction with 
an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
and to quantify these relationships. Prior to constructing the model a natural logarithmic transform was applied
on the variables. This was done so that the coefficients can be interpreted easily as percentage changes.

@@

The \texttt{{bounds\_f\_test}} function in 
the \texttt{ARDL} library, based on the bounds test approach 
outlined in \citeboth{pesaran2001}, was used to calculate the 
F-statistic of the ARDL model and then test for cointegration.

@@

for all $i = 1,2,3,4$.

To confirm that the model did not display autocorrelation, heteroskedasticity, {non-normality}, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively

\subsection{RBF Neural Network}

This study employed radial basis function neural networks, 
a type of artificial neural network that utilise radial basis functions 
at each node in the activation layer. 
This was because radial basis function neural networks 
are well-suited for identifying complex, non-linear relationships 
between time-series variables, making them particularly valuable in 
forecasting tasks. In addition, radial basis function neural networks
have been shown to be effective in operating on small sample sizes, 
which was critical for the data constraints of this study.

\subsubsection{Network Architecture}
The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
a $c$-dimensional target space $\boldsymbol{\mathrm{y}}$. The dataset consists of $N$ input
vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
values $y_n$. The objective is to find a function
  $h(\boldsymbol{\mathrm{x}})$ such that:

@@

By introducing several modifications to the exact interpolation 
procedure, the radial basis function neural network
model is obtained, as presented in Figure \ref{fig:rbf_nn_architecture}. 
The network mapping takes the following form:

@@

    \end{tikzpicture}
    \caption{Radial basis function neural network}
    \label{fig:rbf_nn_architecture}
\end{figure}

@@

The radial basis function neural network was programmed in Python. The Bayesian kernel regression approach outlined by 
\citeboth{bishop1995}, using the \texttt{GaussianProcessRegressor} package in Python was employed to model the network\footnote{See \hyperref[sec: appendix]{Appendix}}.

This approach was then used to conduct the train-test process as follows:
\begin{enumerate}
    \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
    were computed using GridSearch Cross Validation. These hyperparameters were:
    \begin{enumerate}
        \item $\alpha$: a regularisation parameter representing a small quantity
        that was added to the diagonals of the covariance matrix to ensure it was 
        invertible. 
        \item $\beta$: the power in the covariance function, which controlled the shape of the function.
        \item $\sigma$: a parameter which controlled how much of the noise in the data the kernel accounted for. It moderated the risk of overfitting and underfitting.
        \item Length scale: a parameter that controlled the influence that each kernel was to have. Analogous to 
        the width that each radial basis function neuron possesses.
    \end{enumerate}
    \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) were successively passed through the 
    covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
    all training observations, $K$.
    \item A $4$-dimensional input space $X$ took in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
    calculated the posterior weights distribution. 
    Each possible weight vector $w$ represented a different function, explicitly 
    based on the radial basis function covariance function, that could fit the data. In the 
    hidden layer the expectation of this distribution was then taken, which
    combined all the weights and found the mean, weighted by the 
    respective probabilities. This represented the best functional estimate of the price movement. Based on this estimate 
    the predicted output was generated.
    \item The testing process was conducted similarly. The 
    posterior weight distribution that contained the information of the training phase became the new 
    prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
    were calculated, and then the predictive output was found by taking the expectation (\citeboth{rasmussen2006}).

@@

Stock markets provide important information on the economy, 
and economic policymakers are vigilant in tracking the fluctuations 
in these markets, and try to forecast the 
forthcoming fluctuations based on policy changes that affect the macroeconomy.
Within this context, this study examined the 
impacts of the interest rate, USD/GBP
exchange rate, the Consumer Price Index, and 
M3 money supply market activity on
the FTSE250 over the October 1992 – March 2024 period.
The results suggested a cointegrated long-run
relationship between the macroeconomic variables and the stock market index. 
The long-run coefficients implied that
increases in the interest rate and depreciation in the exchange rate
significantly depress the FTSE250 value. 
Policymakers dealing with forecasting the UK stock markets
should consider the factors in the study, and to raise the stock
market's value, a lower interest rate regime should be adopted. This will 
stimulate consumer activity, reduce operating costs for firms, and make 
stocks more attractive investments, thereby increasing the stock market value. 
In addition, the Pound should be increased in value to lower import costs 
and increase the real value of firm earnings, both of which should drive up 
the value of the stock market. The combined effect of these changes will 
create a more stimulating business environment by increasing the capital 
available to firms for research and development, and more efficient 
capital and labour, in turn 
driving economic growth and prosperity. However, 
as with all economic policy shifts, these changes may have 
unintended consequences on other areas of the economy, potentially 
undermining any positive changes. Any policy changes must cautiously be made 
with this at the forefront.

The study also investigated the efficacy of machine learning methods in 
the field of stock-econometric forecasting. The radial basis function network performed well in
short-term forecasting of the FTSE250, and offers investors with an accurate forecasting tool, 
to leverage in their trading strategies. The long-term forecast was able to improve on the ARDL-ECM model, and provides 
economists with a tool to compute non-econometric, quantitative measures of the 
long-term influence of macroeconomic variables on the stock market.

Further research should endeavour to include other macroeconomic variables such as 
foreign direct and portfolio investment, commodity prices and economic growth.
An autoregressive neural network could also be tried, including the past values of the 
FTSE250 in the input space to provide a more accurate model. In addition, 
a more granular data set could be utilised by applying a 
Brownian bridge algorithm to generate daily macroeconomic data. This would allow 
the testing of the radial basis function neural network against other neural networks 
such as the LSTM. Other machine learning methods suited for 
forecasting such as Gradient Boosting Machines could 
also be investigated in future work in this problem domain.

@@

\section*{Appendix}
\label{sec: appendix}

@@

\begin{align} 
    f(x) = x^{T}w, \quad{y = f(x) + \varepsilon}
\end{align}

where $y$ differs from the function values $f(x)$ by noise which is assumed
to follow an iid Gaussian distribution 
$\mathcal{N}_1$ (\citeboth{rasmussen2006}).
\begin{align*}
    \varepsilon \sim \mathcal{N}_1 \bigl(0, \sigma_{n}^2\bigr)
\end{align*}
This assumption together with the model produces the 
probability density of the observations given the inputs, 
which is factored over cases in the training set. This is
permissible because of the independence assumption (\citeboth{rasmussen2006}).

\begin{align}
    p(y|X,w) = \prod_{i=1}^{n} p(y_i|x_i ,w) = \mathcal{N}\bigl(X^T, \sigma_{n}^2 I\bigr)
\end{align}

The weights \( w \) represent the parameters that define the relationship between the inputs and outputs. 
They are treated as random variables defined by the Gaussian distribution \eqref{eq: weight}.

@@

\begin{align}
    k(x,x') = \biggl( ||x-x'|| + \sigma^2\biggr)^{-\beta} \label{eq: gimq}
\end{align}
and used to construct the covariance matrix $K$, which along with 
a mean function $m(x)$ forms the distribution of the the weights

\begin{align}
    w \sim \mathcal{N}\bigl(m(x), K\bigr) \label{eq: weight}
\end{align}

The mean function $m(x)$ acts as a bias, 
providing a baseline expectation of the function values. 
This forms the prior distribution over the weights \eqref{eq: weight}.

@@

\begin{align}
    p(w|y, X) = \frac{p(y|X,w) p(w)}{p(y|X)} \label{eq: post}
\end{align}
Using this, the posterior distribution \eqref{eq:nine} is obtained (\citeboth{rasmussen2006}).

\begin{align}
    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
\end{align}

>>>>

\title{{Investigating the Impact of Macroeconomic Indicators
on the FTSE 250 Index Using Statistical and Machine Learning Methods}}
\author{
  \textbf{Mateen Khan} \\
  University of Warwick  \\ 
  \\
  \textbf{Supervisor: Dr. Randa Herzallah} \\
  Mathematics Institute, University of Warwick
}
\begin{document}

@@

\begin{abstract}
    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate, and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
    In addition, it assesses the effectiveness of radial basis function neural networks (RBFNN) as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables. This study finds quantitative estimates for the short-term and long-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index. It also demonstrates the effectiveness of RBFNN as a short-term forecasting tool for the price of the FTSE 250 Index.
    \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, radial basis function neural network, time-series forecasting}
\end{abstract}

@@

The stock market is a significant component of the financial system and 
plays a crucial role in the broader economy. Consequently, economic policy 
makers take the stock market into account when making fiscal policy 
decisions using a combination of quantitative and qualitative methods (\citeboth{DemirgucKunt1996}). 
These methods such as econometric modelling and forecasting, and empirical 
analysis allow them to understand the short-term and long-term effects of 
macroeconomic indicators
on the wider economy through the stock market.

@@

The significance of this study is to quantify the effects of macroeconomic indicators on the stock market as well as assessing the use of these indicators to forecast stock market prices over the short-term and long-term. In addition, this study seeks to understand which macroeconomic indicators are most relevant in capturing the dynamics of a stock market to aid forecasting.

@@

In order to analyse the dynamic relationship between the variables in the time series data, this study
will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error
Correction Model (ECM). This research also employs an RBFNN to predict the price movements of the FTSE250.

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space and explains the rationale behind the selection of macroeconomic variables.
Section \ref{sec:meth} presents the methodology of this study. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
this study.

@@

RBFNN are a type of artificial 
neural network that use radial basis functions at each node in the 
activation layer of the network.
The advantage of RBFNN lies in their ability to 
operate effectively in volatile financial time-series 
environments (\citeboth{cafferata2019}), their strong ability to 
generalise on such data (\citeboth{sharkawy2020}) as well as their efficacy on small datasets (\citeboth{kosarac2022}). 
\citeboth{esfandyari2016}
supports this, using a sample size of 70 to forecast daily NASDAQ 
Index values using an Artificial Neural Network; they achieved a minimum
test $R^2$ of 83$\%$.

RBFNN have been used considerably for stock market forecasting. \citeboth{cao2004} 
used a RBFNN based on the Optimal Partition Algorithm (OPA) to 
forecast S$\&$P 500 prices. \citeboth{dass2019} use a RBFNN, utilising the Back-propagation 
Algorithm (BPA) instead to tune parameters; their results demonstrate 
that the RBFNN outperformed a standard feed-forward neural network. 
\citeboth{ji2014} also found that the RBFNN was able to predict the value 
of the Shanghai Composite Index on a 45 day time-horizon using the closing price, with a mean prediction error of 
1.4387$\%$. Recently, \citeboth{abotaleb2024}
used quarterly data from 2001 to 2023 to model the influence of the economic 
uncertainty index on the quarterly prices of various stock markets including the 
FTSE 100 Index using a radial basis function. They found that the model 
achieved a relative error of 27.2$\%$ in the case of the FTSE 100 Index.

Whilst \citeboth{abotaleb2024}, used quarterly data and a relatively minor economic indicator, the effectiveness of RBFNN in stock market forecasting using 
major macroeconomic indicators and coarse-grain time-series has not yet been investigated and presents an area for further 
research.

@@

Over the period covered by the dataset, the FTSE250 
showed substantial growth, reflecting the structural transition of the UK economy.
It has also experienced significant volatility, highlighting significant market shifts driven by 
various crises such as the Dotcom bubble burst, and geopolitical shocks.

@@

The USD/GBP exchange rate exhibited moderate variability, highlighting the stable relations between 
the two countries. Nonetheless, geopolitical shocks such as the 2008 financial crisis and Brexit both greatly affected the value of the Pound, 
which accounted for the large range.

The CPI, with an average of 88.36, 
experienced significant fluctuations, which highlighted periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
inflation such as the period between the early to late 2010s.

Interest rates ranged from $0.1\%$ to $8.4\%$ with a standard deviation of $2.52\%$, which reflected the 
increasing role played by monetary policy in the UK since May 1997. It was at this time when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}). 
The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions. For example, 
it made steep rate cuts during the 2008 financial crisis
to stimulate the economy and a series gradual rate hikes in response to 
post-COVID inflationary pressures.

The M3 money supply showed substantial variability, ranging from 494,181 to 3,704,873, 
in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
in response to the recessions caused by the 2008 financial crisis and the COVID-19 pandemic. These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.

@@

This study used an autoregressive distributed lag (ARDL) model in conjunction with 
an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
and to quantify these relationships. Prior to constructing the model,
a natural logarithmic transform was applied
on the variables to enable the coefficients to be interpreted 
easily as percentage changes.

@@

The \texttt{{bounds\_f\_test}} function in 
the \texttt{ARDL} library, which is based on the bounds test approach 
outlined in \citeboth{pesaran2001}, was used to calculate the 
F-statistic and then test for cointegration. The hypotheses were:

@@

where $\beta_{1\leq i\leq4}$ represented the long-term coefficients in 
the ARDL model.

To confirm that the model did not display autocorrelation, heteroskedasticity, {non-normality}, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted respectively

This study employed RBFNN, 
a type of artificial neural network that utilise radial basis functions 
at each node in the activation layer. 
The primary reason for this is that they
are well-suited for identifying complex, non-linear relationships 
between time-series variables, making them particularly valuable in 
forecasting tasks. In addition, RBFNN
have been shown to be effective in operating on small sample sizes, 
which was a significant issue in this study.

\subsubsection{Network Architecture}
The exact interpolation problem, which forms the motivation for RBFNN, 
involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
a $c$-dimensional target space $\boldsymbol{\mathrm{y}}$. The dataset consists of $N$ input
vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
values $y_n$. The objective is to find a function
  $h(\boldsymbol{\mathrm{x}})$ such that:

@@

By introducing several modifications to the exact interpolation 
procedure, the RBFNN
model as shown in Figure \ref{fig:rbf_nn_architecture} is obtained.

The network mapping takes the following form:

@@

    \end{tikzpicture}
    \caption{RBFNN Architecture}
    \label{fig:rbf_nn_architecture}
\end{figure}

@@

The RBFNN was programmed in Python. The Bayesian kernel regression approach outlined by 
\citeboth{bishop1995}, using the \texttt{GaussianProcessRegressor} 
package in the \texttt{sci-kit} library was employed to construct the
network\footnote{See \hyperref[sec: appendix]{Appendix}}.

\subsubsection{Network Training and Testing}

The Bayesian kernel regression 
framework was then used to conduct the train-test process as follows:
\begin{enumerate}
    \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
    were computed using GridSearch Cross Validation. These hyperparameters were:
    \begin{enumerate}
        \item $\alpha$: a regularisation parameter representing a small quantity
        that was added to the diagonals of the covariance matrix to ensure it was 
        invertible. 
        \item $\beta$: the power in the covariance function, which controlled the shape of the function.
        \item $\sigma$: a parameter which controlled how much of the noise in the data the kernel accounted for. It moderated the risk of overfitting and underfitting.
        \item Length scale: a parameter that controlled the influence that each kernel was to have. Analogous to 
        the width that each radial basis function neuron possesses.
    \end{enumerate}
    \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) were successively passed through the 
    covariance function $(x,x')$ to obtain the matrix of the pairwise covariance of 
    all training observations, $K$.
    \item A $4$-dimensional input space $X$ took in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
    calculated the posterior weights distribution. 
    Each possible weight vector $w$ represented a different function, explicitly 
    based on the radial basis function covariance function, that could fit the data. In the 
    hidden layer the expectation of this distribution was then taken, which
    combined all the weights and found the mean, weighted by the 
    respective probabilities. This represented the best functional estimate of the price movement. Based on this estimate 
    the predicted output was generated.
    \item The testing process was conducted similarly. The 
    posterior weight distribution that contained the information of the training phase became the new 
    prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
    were calculated, and then the predictive output was found by taking the expectation (\citeboth{rasmussen2006}).

@@

Stock markets provide important information on the economy, 
and economic policymakers are vigilant in tracking the fluctuations 
in these markets, and try to forecast the 
forthcoming fluctuations based on policy changes that affect the macroeconomy.
Within this context, this study examined the 
impacts of the interest rate, USD/GBP
exchange rate, the Consumer Price Index, and 
M3 money supply market activity on
the FTSE250 between October 1992 and March 2024.
The results suggested a cointegrated long-run
relationship between the macroeconomic variables and the stock market index. 
The long-run coefficients implied that
increases in the interest rate and depreciation in the exchange rate
significantly depress the FTSE250 value. 
Policymakers dealing with forecasting the UK stock markets
should consider the factors in this study, and in order to raise the stock
market's value, a lower interest rate regime should be adopted. This will 
stimulate consumer activity, reduce operating costs for firms, and make 
stocks more attractive investments, thereby increasing the stock market value. 
In addition, the Pound should be increased in value to lower import costs 
and increase the real value of firm earnings, both of which should drive up 
the value of the stock market. The combined effect of these changes will 
create a more stimulating business environment by increasing the capital 
available to firms for research and development, and more efficient 
capital and labour, in turn 
driving economic growth and prosperity. However, 
as with all economic policy shifts, these changes may have 
unintended consequences on other areas of the economy, potentially 
undermining any positive changes. Any policy changes must cautiously be made 
with this at the forefront.

This study also investigated the efficacy of machine learning methods in 
the field of stock-econometric forecasting. The RBFNN performed well in
short-term forecasting of the FTSE250, and offers investors with an accurate forecasting tool, 
to leverage in their trading strategies. The long-term forecast was able to improve on the ARDL-ECM model, and provides 
economists with a tool to compute non-econometric, quantitative measures of the 
long-term influence of macroeconomic variables on the stock market.

Further research should endeavour to include other macroeconomic variables such as 
foreign direct and portfolio investment, commodity prices and economic growth.
An autoregressive neural network could also be tried, including the past values of the 
FTSE250 in the input space to provide a more accurate model. In addition, 
a more granular data set could be utilised by applying a 
Brownian bridge algorithm to generate daily macroeconomic data. This would allow 
the testing of the RBFNN against other neural networks 
such as the LSTM. Other machine learning methods suited for 
forecasting such as Gradient Boosting Machines could 
also be investigated in future work in this problem domain.

@@

\section{Appendix}
\label{sec: appendix}

@@

\begin{align} 
    f(\boldsymbol{\mathrm{x}}) = \boldsymbol{\mathrm{x}}^{T}\boldsymbol{\mathrm{w}}, \quad{\boldsymbol{\mathrm{y}} = f(\boldsymbol{\mathrm{x}}) + \varepsilon}
\end{align}

where $\boldsymbol{\mathrm{y}}$ differs from the function values $f(x)$ by noise which is assumed
to follow an iid Gaussian distribution 
$\mathcal{N}_1$ (\citeboth{rasmussen2006}).
\begin{align*}
    \varepsilon \sim \mathcal{N}_1 \bigl(0, \sigma_{n}^2\bigr)
\end{align*}
This assumption together with the model produces the 
probability density of the observations given the inputs, 
which is factored over cases in the training set. This is
permissible because of the independence assumption (\citeboth{rasmussen2006}).

\begin{align}
    p(\boldsymbol{\mathrm{y}}|X,\boldsymbol{\mathrm{w}}) = \prod_{i=1}^{n} p(\boldsymbol{\mathrm{y}}_i|\boldsymbol{\mathrm{x}}_i ,\boldsymbol{\mathrm{w}}) = \mathcal{N}\bigl(X^T, \sigma_{n}^2 I\bigr)
\end{align}

The weights \( \boldsymbol{\mathrm{w}} \) represent the parameters that define the relationship between the inputs and outputs. 
They are treated as random variables defined by the Gaussian distribution \eqref{eq: weight}.

@@

\begin{align}
    k(\boldsymbol{\mathrm{x}},\boldsymbol{\mathrm{x}}') = \biggl( ||\boldsymbol{\mathrm{x}}-\boldsymbol{\mathrm{x}}'|| + \sigma^2\biggr)^{-\beta} \label{eq: gimq}
\end{align}
and used to construct the covariance matrix $K$, which along with 
a mean function $m(\boldsymbol{\mathrm{x}})$ forms the distribution of the the weights

\begin{align}
    \boldsymbol{\mathrm{w}} \sim \mathcal{N}\bigl(m(\boldsymbol{\mathrm{x}}), K\bigr) \label{eq: weight}
\end{align}

The mean function $m(\boldsymbol{\mathrm{x}})$ acts as a bias, 
providing a baseline expectation of the function values. 
This forms the prior distribution over the weights \eqref{eq: weight}.

@@

\begin{align}
    p(\boldsymbol{\mathrm{w}}|\boldsymbol{\mathrm{y}}, X) = \frac{p(\boldsymbol{\mathrm{y}}|X,\boldsymbol{\mathrm{w}}) p(\boldsymbol{\mathrm{w}})}{p(\boldsymbol{\mathrm{y}}|X)} \label{eq: post}
\end{align}
Using this, the posterior distribution \eqref{eq:nine} is obtained (\citeboth{rasmussen2006}).

\begin{align}
    p(\boldsymbol{\mathrm{w}}|\boldsymbol{\mathrm{y}}, X) \sim \mathcal{N}\bigl(\frac{A^{-1}X\boldsymbol{\mathrm{y}}}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
\end{align}
